---
title: "Initial Endpoint Assessment with R Markdown & Python"
runtime: shiny
output: 
  html_document:
    code_folding: show
    theme:
      color-contrast-warnings: false
      bg: "#202123"
      fg: "#B8BCC2"
      primary: "#EA80FC"
      secondary: "#00DAC6"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup, include=FALSE}
library(reticulate)
bslib::bs_themer()
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")

```

## On-Prem Storage Analyzer

The following code chunk focuses on importing and executing a separate python script that provides a bar graph to visualize the storage usage on all mapped storage drives on the host machine it's executed on.


```{python, Storage-Analyzer}
import matplotlib.pyplot as plt
import numpy
import psutil

# The following function searches for available mounted drives and obtains their usage information and stores the data in a dictionary. Finally, the details are then printed.

def disk_use():
    parts = psutil.disk_partitions(all=True)
    storage_dict = {}

    for partition in parts:
        try:
            usage = psutil.disk_usage(partition.mountpoint)
            storage_dict[partition.mountpoint] = {
                'Total': usage.total,
                'Used': usage.used,
                'Free': usage.free,
                'Percentage': usage.percent
                }
        except Exception as ex:
            print(f"Error retrieving disk usage for {partition.mountpoint} {ex}")
        
    return storage_dict
    
storage_dict = disk_use()
for mountpoint, info in storage_dict.items():
        print(f"Mountpoint: {mountpoint}")
        print(f"Percentage: {info['Percentage']}%")
        print("-" * 50)

# The following next function utilizes the previously defined dictionary now populated with data to usage with a bar craph to easily visualize these percentages of utilization across all mounted drives on the machine.

def graph(storage_dict):
     mountpoints = list(storage_dict.keys())
     Percentages = [info['Percentage'] for info in storage_dict.values()]

     plt.style.use('ggplot')
     x = numpy.arange(len(mountpoints))
     y = Percentages  

     fig, ax = plt.subplots()

     ax.bar(x, y, width=1, edgecolor="grey", linewidth=0.7)
     ax.set_xticks(x)
     ax.set_xticklabels(mountpoints, rotation=0, ha='right')

    # Set y-axis labels and ticks
     ax.set_yticks(numpy.arange(0, max(Percentages) + 1, 10))

     ax.set_ylabel('Percentage (%)')
     ax.set_title('Disk Usage Percentage')

     plt.tight_layout()
     plt.show()

graph(storage_dict)
```

## Network interfaces:

This next section is designed to provide a simple output on the network interfaces detected on the host machine.


```{r, net-interfaces}
library(processx)
library(knitr)

net_interfaces <- processx::run("ipconfig")
# Split the output into lines
output_lines <- strsplit(net_interfaces$stdout, "\n")[[1]]

# Print each line
for (line in output_lines) {
  print(line)
}

```


## Local User Enumeration:

This next section provide some basic enumeration on the host machine and provides the raw output of all local and domain information on the machine.:

```{r, local-user-information}

library(processx)
library(knitr)

local_user_enum <- processx::run("whoami", args = "/all") 

# Split the output into lines
output_lines <- strsplit(local_user_enum$stdout, "\n")[[1]]

# Print each line
for (line in output_lines) {
  print(line)
}

```

## Further AD Involvement:

This next code chunk focuses primarily on enumerating the machine for domain user access.

##### Note: This code chunk is disabled by default. Uncomment each line to enable this code chunk and rerun it to execute the script.

```{r, ad-information}

#dom_user_enum <- system2("net", args = c("user", "/domain"))

# Split the output into lines
#output_lines <- strsplit(dom_user_enum$stdout, "\n")[[1]]

# Print each line
# for (line in output_lines) {
#  print(line)
#}

```


## System Information

Next, the following code chunk focuses on displaying the output of systeminfo to see the current current system details at a glance.

```{r, sysinfo}

sys_info <- processx::run("systeminfo")

# Split the output into lines
output_lines <- strsplit(sys_info$stdout, "\n" [[1]])

# print each line
for (line in output_lines) {
  print(line)
}
```























